Topic modeling bla bla

### 2017-06-06 thesis_report/draft_thesis.tex
Fix: variational inference

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{biblatex}
\addbibresource{sample.bib}

\begin{document}
\maketitle
\tableofcontents

@@

We can analyze a corpus of documents with LDA by examining the posterior 
distribution of topics $\beta$, topic proportions $\theta$ and topic 
assignments $z$ conditioned on the documents. This reveals latent structure 
in the collection that can be used for prediction or data exploration. This 
posterior can't be computed directly; it is usually approximated using Markov 
Chain Monte Carlo (MCMC) or variational inference. Both methods are effective, 
but also present significant computational challenges in large data sets \cite{onlineLDAvb}. In our implementation, we're using the online variational 
LDA model, which is described in section 2.2.3.4.

@@

Variational inference can be seen as an extension of the EM algorithm and is
used to approximate intractable integrals. The goal of variational inference is
to approximate an intractable probability distribution $p$, with a tractable
one $q$, in a way that makes them as close as possible given a similarity
measure. In this thesis, variational inference is used to approximate the
posterior distribution of latent variables.

For a bayesian model where x = x 1:n are observations and z = z 1:m are latent
variables, the posterior distribution is

\begin{equation}
p(x \mid z) = \frac{p(x \mid z) p(z)}{\int\limits_{z} p(x \mid z)p(z)}
\end{equation}

For many models, this distribution is intractable due to the integral in the
denominator. The posterior can be computed up to a constant, but the
normalization constant cannot be computed. Variational inference methods can be
used to approximate this integral. Applying variational inference to posterior
distributions is sometimes called variational Bayes but is only an instance of
a more general problem that variational inference solves.

@@

>>>>

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{algorithm2e}

\begin{document}

\maketitle
\tableofcontents

@@

We can analyze a corpus of documents with LDA by examining the posterior
distribution of topics $\beta$, topic proportions $\theta$ and topic
assignments $z$ conditioned on the documents. This reveals latent structure in
the collection that can be used for prediction or data exploration. This
posterior can't be computed directly; it is usually approximated using Markov
Chain Monte Carlo (MCMC) or variational inference. Both methods are effective,
but also present significant computational challenges in large data sets
~\cite{onlineLDAvb}. In our implementation, we're using the online variational
LDA model, which is described in section 2.2.3.4.

@@

Learning the various distributions (the set of topics, their associated word
probabilities, the topic of each word, and the particular topic mixture of each
document) is a problem of Bayesian inference. In the original paper it's being
used a variational Bayes approximation of the posterior distribution;
~\cite{blei2003latent} alternative inference techniques use Gibbs sampling and
expectation propagation. The goal of variational inference is to approximate an
intractable probability distribution $p$, with a tractable one $q$, in a way
that makes them as close as possible given a similarity measure. In this thesis,
variational inference is used to approximate the posterior distribution of
latent variables. A new distribution over the hidden variables $q(z, \beta)$
(called the variational distribution) is  introduced. This new distribution has
properties such that it can be efficiently computed.

The variational distribution is a function of a set of free parameters that are optimized such that the variational distribution is as close as possible to the 
actual target posterior distribution where closeness is measured in terms of KL divergence.
Minimizing the KL divergence between the variational distribution and the target
posterior is equivalent to maximizing the Evidence Lower BOund (ELBO) which is

\begin{equation}
L(q) = E_q [log p(x, z, \beta)] -  E_q [log q(z, \beta)]  
\end{equation}

The variational distribution has the property that it can be efficiently
computed. This is done by making each hidden variable independent of each other.

\begin{equation}
q(z, \beta) = q(\beta \mid \lambda) \prod\limits_{n=1}^N \prod\limits_{j=1}^J q(z_{nj} \mid \phi_{nj})   
\end{equation}

The posterior over the per-word topic assignments $z$ is parameterized by
$\phi$, the posterior over per-document topic weights $\theta$ is parameterized
by $\gamma$, and the posterior over the topics $\beta$ is parameterized by
$\lambda$. Furthermore, each hidden variable is governed by its own variational
parameter so the variational parameters $\lambda$ governs the global variables
$\beta$ and the variational parameters $\phi_n$ govern the local variables $z_n
q(\beta \mid \lambda)$ and $q(z_{nj} \mid \phi_{nj})$ take the same form as the
complete conditionals $p(\beta \mid x, z, \alpha)$ and $p(z_{nj} \mid x_n, z_{n,
-j}, \beta)$, but the natural parameters are now $\lambda$ and $\phi_{nj}$,
respectively to give

\begin{equation}
q(\beta \mid \lambda) = h(\beta) exp(\lambda^T t(\beta) - \alpha_g(\lambda))
\end{equation}
\begin{equation}
q(z_{nj} \mid \phi_{nj}) = h(z_{nj}) exp(\phi_{nj}^T t(z_{nj}) - \alpha_l(\phi_{nj}))
\end{equation}

Where $h(.)$ is called the base measure, $\alpha(.)$ is called the log-
normalizer, $\eta(.)$ is called the natural parameter, and $t(.)$ is called the
sufficient statistics. Furthermore, we also assume that the prior distribution
over $\beta$ is part of the exponential family. We maximize the ELBO objective
function with a coordinate ascent procedure. We find its gradient with respect
to the global variational parameter $\lambda$ and find the value of $\lambda$
that sets the gradient to zero. We do the same thing for the local parameters
$\phi_n$. We iterate between these updates until we converge to the maximum of
the ELBO. The updates are given without proof, but the general procedure is to
write the ELBO in terms of parameter of interest (either $\lambda$ or $\phi_n$)
then take the gradient and set it to zero.

\begin{equation}
\lambda = E_q[\eta_g(x, z, \alpha)]
\end{equation}
\begin{equation}
\phi_{nj} = E_q[\eta_l(x_n, z{n, -j}, \beta)]
\end{equation}

Therefore the updates of each variational parameter are equal to the expected
value of the natural parameters of the complete conditionals. A simple outline 
of the algorithm ~\cite{alexminnar} is given below

\begin{algorithm}[H] 
\SetAlgoLined

Initialize $\lambda^{(0)}$ randomly.
\Repeat{the ELBO converges}{
	\For{each local variational parameter $\phi_{nj}$}{\label{forins}
		Update $\phi_{nj}^{(t)} = \mathbb{E}_{q^{(t - 1)}}[\eta_{l,j}(x_n, z_{n,-j}, \beta)]$.
	}
	Update the global variational parameters, $\lambda^{(t)} = \mathbb{E}_{q^{(t)}}[\eta_g (z_{1:N}, x_{1:N})]$.
}
\caption{Variational Bayes for LDA} 
\end{algorithm}

As you can see, in the local parameter update (in the $for$ loop), we have to
iterate over every data point in the data set which is computationally
expensive. The algorithm above has constant memory requirements and empirically
converges faster than collapsed Gibbs sampling. However, it still requires a
full pass through the entire corpus in each iteration. It can be slow to apply
to very large datasets. An online variational inference algorithm is proposed by
Hoffman et all. ~\cite{onlineLDAvb} for fitting $\lambda$, the parameters to the
variational posterior over the topic distributions $\beta$. The "online LDA" is
nearly as simple as the variational algorithm, but converges much faster for
large datasets.

@@

\bibliography{references}{}
\bibliographystyle{plain}
